Envelope Estimation using Geometric Properties of a Discrete Real Signal

### 2021-05-08 Papers/01_Pulses_Envelope/00ElsevierPulsesEnvelope.tex
preparing for final revisions

\author{Carlos Henrique Tarjano Santos\corref{cor1}}
\ead{carlostarjano@id.uff.br}
\address{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
\cortext[cor1]{Corresponding author, carlostarjano@id.uff.br}

\author{Valdecy Pereira}
\address{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}

@@

% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
After defining what will be considered an envelope in the context of this work, in this section we explore some simplifications that can be applied to a discrete wave, defining the representation that will be used in the subsequent part of the work.

We next explore the concept of discrete curvature, putting forth our own method, tailored to the definition of the equivalent circle of a wave; we end this section by showing how this circle can be used to identify the envelope of a discrete wave, in a similar approach to that of the alpha shapes theory, but taking advantage of some unique structure present in discrete waves.

The rest of this paper is divided as follows: We start by defining as specifically as possible what a temporal envelope is, especially in the context of this work and providing a mathematical definition, in terms of the Hilbert transform, for the case of a narrowband signal.

We proceed to illustrate the representation of a discrete wave used in this work, largely based on the concept of dividing the wave into its constituent pulses, and how it can simplify the algorithm, reducing the high dimensionality generally present in digital representations of waves.

Next, we review methods for the estimation of the curvature of a discrete wave, after which we present our own; this method will then be used to estimate the equivalent circle, that will be used to define the envelope of the wave, via the procedure described in sequence.

In the results section, we present some illustrations of the envelopes obtained with our method, comparing both accuracy and running time with the most common algorithms encountered in the literature. We also suggest an alternative use of the algorithm to identify the negative and positive frontiers of a wave, that can be seen as partial envelopes considering only the positive and negative pulses of a wave, respectively.

We follow with suggestions about how the approach here presented can be useful beyond envelope detection, exemplifying with some applications, such as the extractions of the average waveform of a digital wave, and concluding with commentary on potential future developments of the method.

Interactive visualizations are available from the repository prepared for this work \cite{2020TarjanoEnvelope}, that also contains an implementation of the method in pure Python.
A Python module available at the Python Package Index (PyPI) installable via the command `pip install signal-envelope` for convenience, is also available. Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module, and can be separately obtained via the release section in the repository.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Temporal envelope}\label{sec:Characterization of a temporal envelope}

@@

From \ref{fig:Envelope} and the discussion in Section \ref{sec:Characterization of a temporal envelope}, we can see that only the maximum point of a positive pulse or the minimum point of a negative pulse can potentially be equal to the envelope and are, thus, our only points of interest. We can then proceed, for the rest of the method, considering only those points, to great computational economy.

@@

The effect of this normalization can be seen in the points shown in Fig.\ref{fig:Coordinates}, shown in scale for both basis. We can consider our points now residing in the Cartesian coordinate system.

@@

\section{Discrete Curvature Estimation}

@@

\section{Extracting the envelope}

@@

\section{Quality of an envelope}\label{sec:quality}

@@

\subsection{Comparison with traditional algorithms}

@@

\section{Extensions}

>>>>

% \author{Carlos Henrique Tarjano Santos\corref{cor1}}
% \author{Valdecy Pereira}
% \address{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}
% \cortext[cor1]{Corresponding author}
% \ead{carlostarjano@id.uff.br}

\cortext[cor1]{Corresponding author}
\address{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}

\author{Carlos Henrique Tarjano Santos\corref{cor1}}
\ead{carlostarjano@id.uff.br}

\author{Valdecy Pereira}
% \ead{support@elsevier.com}

% \author{Valdecy Pereira}
% \address{Department of Production Engineering, Universidade Federal Fluminense, Rua Passo da Pátria, 156, Campus Praia Vermelha, Bloco D - sala 309, São Domingos, Niterói, RJ, Brasil, CEP: 24.210-240}

@@

We start with a brief background of envelope detection in Section \ref{sec:envelope} and define, as specifically as possible, what constitutes a temporal envelope.

In Section \ref{sec:representation} we explore some simplifications to the task of detecting the envelope of a signal that arise from the envelope definition adopted in this work and simplify the task of envelope detection.

Section \ref{sec:mapping} explores the geometric characteristics of envelope detection and formalizes how this classic DSP problem can be translated to the geometric problem of finding the shape of a set of points.

We next explore, in Section \ref{sec:curvature}, the concept of discrete curvature, that will be used in Section \ref{sec:envelopedetection}, where the envelope detection algorithm is presented.

Section \ref{sec:quality} discusses the difficulties in the assessment of the quality of an envelope, proposing qualitative and quantitative approaches, and demonstrates that the method propose conforms to four requirements for a physically plausible envelope proposed in \cite{1996Loughlinamplitude}.

In Section \ref{sec:comparison} we compare the presented algorithm with traditional envelope detection methods, both in terms of accuracy and time and present, in Section \ref{sec:extensions},  potential applications of the algorithm, some of which might not be immediately obvious.

% TODO %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Interactive visualizations are available from the repository prepared for this work \cite{2020TarjanoEnvelope}, that also contains an implementation of the method in pure Python.
A Python module available at the Python Package Index (PyPI) installable via the command `pip install signal-envelope` for convenience, is also available. Besides, those working under Windows 64-bit can benefit from a specialized dynamic library coded in C++ that is automatically used within this module, and can be separately obtained via the release section in the repository.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section{Temporal envelope}\label{sec:envelope}

@@

From \ref{fig:Envelope} and the discussion in Section \ref{sec:envelope}, we can see that only the maximum point of a positive pulse or the minimum point of a negative pulse can potentially be equal to the envelope and are, thus, our only points of interest. We can then proceed, for the rest of the method, considering only those points, to great computational economy.

@@

The effect of this normalization can be seen in the points shown in Fig. \ref{fig:Coordinates}, shown in scale for both basis. We can consider our points now residing in the Cartesian coordinate system.

@@

\section{Discrete Curvature Estimation} \label{sec:curvature}

@@

\section{Identifying the envelope} \label{sec:envelopedetection}

@@

\section{Quality of an envelope} \label{sec:quality}

@@

\subsection{Comparison with traditional algorithms}\label{sec:comparison}

@@

\section{Extensions} \label{sec:extensions}
